---
title: "HW 1"
author: "Barbara"
date: "November, 2015"
---
###About
This was my first project for the **Reproducible Research** course in Coursera's Data Science specialization track. 

# Reproducible Research: Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

### Loading and preprocessing the data
##### Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
##### Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

-----

### What is mean total number of steps taken per day?
```{r}
stepsPerDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

##### Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsPerDay, xlab='Total steps for each day', ylab='Frequency with binwith 600', binwidth=600)
```

##### Calculate and report the mean and median total number of steps taken per day
```{r}
stepsPerDayMean <- mean(stepsPerDay)
stepsPerDayMedian <- median(stepsPerDay)
```
* The mean is: `r stepsPerDayMean`
* The median is:  `r stepsPerDayMedian`

-----

### What is the average daily activity pattern?
```{r}
averagePerTimeblock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### Make a time series plot
```{r}
ggplot(data=averagePerTimeblock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5 minute interval") +
    ylab("average number of steps") 
```

##### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
moststep <- which.max(averagePerTimeblock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averagePerTimeblock[moststep,'interval'])
```

* Most Steps are onn the interval: `r timeMostSteps`

----

### Imputing missing values
##### Calculate and report the total number of missing values in the dataset 
```{r}
numMissing <- length(which(is.na(activityData$steps)))
```

* The number of missing values is: `r numMissing`

##### Devise a strategy for filling in all of the missing values in the dataset.
##### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityImputed <- activityData
activityImputed$steps <- impute(activityData$steps, fun=mean)
```


##### Make a histogram of the total number of steps taken each day 
```{r}
stepsPerDayImputed <- tapply(activityImputed$steps, activityImputed$date, sum)
qplot(stepsPerDayImputed, xlab='Total steps each day', ylab='Frequency using binwith 600', binwidth=600)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsPerDayMeanImputed <- mean(stepsPerDayImputed)
stepsPerDayMedianImputed <- median(stepsPerDayImputed)
```
* The mean (Imputed) is: `r stepsPerDayMeanImputed`
* The median (Imputed) is:  `r stepsPerDayMedianImputed`


----

### Are there differences in activity patterns between weekdays and weekends?
##### Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.

```{r}
activityImputed$dateType <-  ifelse(as.POSIXlt(activityImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### Make a panel plot containing a time series plot

```{r}
averagedActivityImputed <- aggregate(steps ~ interval + dateType, data=activityImputed, mean)
ggplot(averagedActivityImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
